\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
%\usepackage{fullpage}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{algorithmic}
\usepackage{wrapfig}

\title{CS263: HW \#7}
\author{Ben Lickly}
\date{April 13, 2010}

\newcommand{\problem}[1]
{\subsubsection*{} %\fbox{\parbox{\textwidth}{
\vspace{-16pt} \section{} \vspace{-22pt} \qquad
#1%}}
\bigskip \bigskip
}

\newcommand{\while}[2]{\operatorname{while}\, #1\ \operatorname{do}\ #2}
\newcommand{\ifthen}[3]{\operatorname{if}\, #1\ \operatorname{then}\ #2\ \operatorname{else}\ #3}
\newcommand{\denote}[1]{\llbracket #1 \rrbracket}
\newcommand{\proves}{\vdash}
\newcommand{\axiomatic}[3]{\{#1\}\ #2\ \{#3\}}
\newcommand{\meet}{\bigwedge}
\newcommand{\powerset}{\mathcal{P}}
\newcommand{\glb}{\operatorname{glb}}

\begin{document}
\maketitle

\problem{What is the type of the combinators $K$ and $S$ in the polymorphic
lambda calculus $F_2$ ? Argue that the combinator $D$ does not have a type in
$F_2$.
}

The $K$ combinator is the unique function (in the pure $F_2$
lambda calulus) with type
$\Lambda t_1. \Lambda t_2. t_1 \to t_2 \to t_1$.

The $S$ combinator has type
$\Lambda t_1. \Lambda t_2. \Lambda t_3.
(t_1 \to t_2 \to t_3) \to (t_1 \to t_2) \to t_1 \to t_3$

The $D$ combinator can not have a type because it would need to satisfy the
equation $t(t) = t$.  We may think naively that the type $\Lambda t. t \to t$
would work here, but this allows any $t$, whereas we need to restrict ourselves
only to function types (whose argument and return are both function types whose
argument and return are \dots).  We may then try
$\Lambda t. (t \to t) \to (t \to t)$, but we will run into the same problem one
level deeper.  It turns out that no finite $F_2$ type expression can express
the need that there are functions ``all the way down''.

\problem{What would be a good subtyping rule for continuations?  That is, under
what conditions it is sound to say that $T_1 cont \le T_2 cont$.
}

If a program fragment only uses $callcc$, then its uses of continuations are
covariant and we can say that $T_1 \le T_2 \implies T_1 cont \le T_2 cont$.

If a program fragment only uses $throw$, then its uses of continuations are
contravariant and we can say that $T_2 \le T_1 \implies T_1 cont \le T_2 cont$.

In the general case, however, we can not say that continuations are either
covariant or contravariant.

\problem{
{\bf Theorem 1} For all $e$ and $\tau$ , if $\proves e:\tau$ then
$\proves_0 e:\tau$.
}

We will try to prove this on the structure of the expression, $e$.

The base cases here are when $e$ is a constant or variable.

Base Case 1:  $e$ is a constant.

This case seems to violate the theorem.  For example, we can get
$\proves 1 : real$, but we can't seem to prove that $\proves_0 1 : real$.

Thus, we have found a counterexample that refutes our theorem.  \qed

\medskip

Even if we try to correct (or ignore) this problem with constants, it seems
that the theorem still does not hold in the case of lambda expressions,
since our new rule only applies to variables.

For example, we have $\proves (\lambda x_{:real} . x) : (int \to real)$,
by first proving a type of $real \to real$ and then using subsumption.
We cannot prove the equivalent judgement with our new rule, however:
\[\proves_0 (\lambda x_{:real}. x) : (int \to real)\]
This is not true, since our new rule only applies to variables.

\end{document}
